\theseminartitle \\ \small \thetitle

### 2021-09-27 tex/paper.tex
Minor adaptions to paper

In this paper we give an overview of the general concepts of system benchmarks before we consider in detail how classical benchmarking approaches can be used and extended to integrate it into the standard development process of software. Specifically, we take a look at how benchmark can be used to swiftly detect and react to performance changes throughout the development process. In order to do this we will take a look at Change-Point detection algorithms. We will outline and analyse the methodology and opportunities of such approaches while taking a careful look at the challenges that may arise due to this.

To support and evaluate our claims, we will conduct a practical experiment where we will use and compare several benchmarking techniques on a piece of well known, open-source software.

\section{System Benchmarks}
\label{sec:benchmarking}
The term "Benchmark" is a broad term that can be interpreted and defined in many different ways depending on the context and application domain. We will use this chapter to give a definition of Benchmarking how we will use it throughout this paper.

@@

While Fixed-Work benchmarks are easy to understand and implement, it introduces an implicit performance bottleneck that is also known as \textit{Amdahl's Law}. It states that the maximum possible performance improvement that can be achieved in Fixed-Work Benchmarks is prohibitly upper-bounded by the fraction of work that is performed by the component that is being improved (See \cite{Kounev} p. 9f).

@@

	Both of these challenges are heavily influenced by the length of software release cycles and the resulting amount of commits throughout one cycle. As an example one can take a look at a few open-source repositories. The C++ test framework \textit{googletest}\footnote{See \url{https://github.com/google/googletest}} by Google with it's most recent release (\textit{1.11.0}) in June 2021, while the last release before that (\textit{1.10.0}) was released in October 2019. During this period of 28 months roughly 700 commits have been added to the repository. Similarly for the C implementation of the LZ4 compression algorithm that can be found on GitHub\footnote{\url{https://github.com/lz4/lz4}} in the 15 months between the two most recent releases about 250 commits have been added to the code.

@@

	\subsection{Continuous Benchmarking}
In this section a technique that will further be reffered to as \textit{Continuous Benchmarking} will be elaborated further. First the motivation of this approach will be discussed to give a working definition. Secondly a selection of possible continuous benchmarking strategies will be outlined. Lastly the possible benefits and potential challenges will be showcased and weighted to give an assessment of using such a technique.

	\subsubsection{Definition}
	As previously discussed in \autoref{ssec:bench_rel} benchmarking only before releases introduces several challenges. The general motivation of continuous benchmarking is to address these by, simply said, generally adding more data points to evaluate. These added data points can then be used for a more detailled analysis of the systems performance over time. The general idea is similar to those used in \gls{ci} and \gls{cd}. By running tests and releases more frequently, often after every commit, these methods aim to shorten release cycles. \gls{cb} however, is a term and methodology that seems not yet commonly defined in academia.

@@

	\paragraph{}As for any benchmark also \gls{cb} should only be run against a correctly working software. Therefore if integrated into a contiuous integration pipeline, \gls{cb} should be among the last steps such that the correct functionality of the software is already ensured by the \gls{cd} tests beforehand (See \cite{grambow2019} p. 2).

	Similar to \gls{cd} approaches in \gls{cb} the question also arises when to run the \gls{cb}. For \gls{cd}, optimally all tests and operations should be performed after every commit. However in practice this might not always be feasible due to the amount of tests and the associated compute time with that which is why complex software systems often alrady use a multi-tiered approach to \gls{cb}. A similar question arises when deciding on when to run \gls{cb}. Running benchmarks after every commit can pose challenges which will be further evaulated in the following section. However for this paper our main focus will not be on when to run \gls{cb} but rather how the data produced by a \gls{cb} can be used to detect performance changes.

@@

		Using \gls{cb} opens up several opportunities for improvements to the development process of software. First of all it gives a regular overview of a software systems performance. However the usability of these results is also heavily dependent on the quality of your testbed. Especially the use of cloud infrastructure can introduce additional noise to the performance of a system. As has been shown by Daly and Ingo in \cite{daly2019} reducing the noise in such a scenario is a heavily application specific task that can take several weeks of investigating the systems behavior.

	Running performance benchmarks more frequently naturally also enables to detect performance changes more quickly. As a result it also makes it easier to isolate the cause of the change as there may only be fa few commits between the two benchmarks. However again the comparability of these results heavily depends how and if we can take into account the testbeds noise accordingly as to not errornously flag outliers from noise as actual performance changes. In \cite{daly2021} Daly proposed a technique called \textit{canary testing} to counter these issues. However, in their evaluation canary testing was not viable in their productive system and resulted in more cost than benefit (See \cite{daly2021} p. 6f).

	Another point to consider when using \gls{cb} is the additional data that will inherently get produced. Especially for modern software systems there is not only a single benchmark that is of interest but rather multiple benchmarks with for example different system configurations or workloads. Therefore as Daly et al. and Grambow et al. conclude in their works that manually reviewing the performance data for \gls{cb} is not feasible(See \cite{daly2020} p. 2, \cite{grambow2019} p. 5). As a result, to assist in \gls{cb} and especially in detecting performance changes, automated approaches and toolings are required to preprocess this data and possibly detect performance changes automatically. Approaches for such tools have been described by Daly et al. (\cite{daly2020} p. 2, \cite{daly2021} p. 2ff) and Grambow et al.(\cite{grambow2019}). One of the techniques commonly referred to in this context is change point detection which will be covered in a little more detail in \autoref{sec:cp_detection}.

@@

	\section{Change Point Detection}
	\label{sec:cp_detection}
	This chapter will give a brief introduction into a technique from signal processing known as \textit{Changepoint Detection}. It will briefly introduce the term Change point and it's distinction from outliers adn introduce one specific change point detection algorithm.

@@

	\begin{figure}[ht!]
		\centering
		\begin{subfigure}[b]{0.45\textwidth}
			\centering
			\includegraphics[width=\textwidth]{outlier}
			\caption{Visualization of an outlier}
		\end{subfigure}
		\begin{subfigure}[b]{0.45\textwidth}
			\centering
			\includegraphics[width=\textwidth]{changepoint}
			\caption{Visualization of a change point}
		\end{subfigure}
		\caption{Visual difference of an outlier and a changepoint}
		\label{fig:outlier_cp}
	\end{figure}

	In other words, a change point indicates that the overall underlying signal undergoes a noticable change after that point, while an outlier is only a temporary effect that generally does not persist. \autoref{fig:outlier_cp} shows a visualization of this distinction.

	The challenge of finding such change points is commonly referred to as "Change Point Detection" and is defined by Aminikhanghahi and Cook as "the problem of finding abrupt changes in data when a property of the time series changes" (\cite{Samaneh2016}). In the domain of (continuous) benchmarking, change point detection can be used as a tool to determine points at which a performance regression (or improvement) occurs. With this regard, the performance of a system throughout different commits can be interpreted as a series of time-ordered series, which represents the underlying signal that is analyzed by a chnge-point detection algorithm (See \cite{daly2020} p. 69)

	\subsection{E-Divisive Means}
	\label{ssec:cp_edivisive}
	One algorithm for detecting change points is the E-Divisive Means algorithm. In contrast to other change point detection algorithms it offers a few benefits decribed by Daly et al. (See \cite{daly2020} p. 69):

	\begin{itemize}
		\item No distributional assumptions required
		\item Can identify multiple change points using recursion
		\item Does noit require training data
	\end{itemize}

	The algorithm is based on hierachically dividing the given time-series into multiple clusters. The recursive approach can be summarized in three steps:
	\begin{enumerate}
		\item For each point in the time series calculate a statistic $\hat{Q}$
		\item Select the largest $\hat{Q}$ as the next change point
		\item This divides the time series into two new clusters
	\end{enumerate}

@@

	To perform the experiment as described in \autoref{ssec:exp_goals} the open-source software LZ4\footnote{\url{https://github.com/lz4/lz4}} will be used. \texttt{LZ4} is a compression library with an active commit history. In total \texttt{LZ4} offers 31 releases and more than 2500 commits in total\footnote{As of 16.09.2021}.

@@

	Lzbench offers the option to disable all the other compression algorithms at compile time. The LZ4 version that is used for the benchmark can simply be changed by overwriting the corresponding lz4 files within the lzbench repository before compilation. In the experiment a slightly adapted version of Lzbench will be used that has the other compression algorithms disabled at compile time and also incorporates some minor changes that were required to ensure compilation on the cluster infrastructure.

@@

Since the LZ4 repository consists of over 2500 commits benchmarking each individual commit is infeasible for the scope of this paper as it would both generate too much data to evaluate properly as well as strain the available shared server resources too much. Therefore a two staged approach will be used:
\begin{itemize}
	\item In the first step, a benchmark will be performed on the publicly available releases\footnote{Based on the tags created in the GitHub repository}. The results of these benchmarks will be used to identify a commit range with a significant performance change
	\item All commits in the identified commit range will be benchmarked for a more detailed and thorough analysis.
\end{itemize}

The benchmark itself will be performed on the Cluster Infrastructure of the Software Engineering Chair of Saarland University using Slurm\footnote{See \url{https://slurm.schedmd.com/documentation.html}}. The machines that are used are part of the \texttt{kine} cluster with the following hardware configuration\footnote{CPU Information has been acquired using the \texttt{lscpu} command}:

@@

	Based on these results the commit range between \textit{v1.8.1} and \textit{v1.8.2} was selected as the commit range for a more thorough and detailled analysis. The results of this analysis are further discussed in \autoref{sec:exp_evaluation}. However as the data for the lz4 benchmark only indicated a performance improvement and the goal of this work is to evaluate the detection both performance improvements and regressions, the release benchmarks were repeated with the \texttt{lz4hc} compression algorithm. The \texttt{lz4hc} implementation is a variant of the \texttt{lz4} implementation that features a higher compression rate at the cost of lower compression speed. As \texttt{lz4hc} is also included in the same source code as \texttt{lz4} the only adaption required to the benchmark was to replace the compression algorithm used in the command line parameters passed.

>>>>

This paper gives an overview of the general concepts of system benchmarks and describes how classical benchmarking approaches can be used and extended to integrate them into the standard development process of software. Specifically, it will outline how benchmarks can be used to swiftly detect and react to performance changes throughout the development process. In order to achieve this the paper will describe and use the concepts of change point detection. This paper will outline and analyse the methodology and opportunities of such approaches while taking a careful look at the challenges that may arise due to this.

To support and evaluate the claims, this paper includes a practical experiment that will serve as a basis to compare two benchmarking approaches on a piece of well known, open-source software.

\section{System Benchmarks}
\label{sec:benchmarking}
The term "Benchmark" is a broad term that can be interpreted and defined in many different ways depending on the context and application domain. This chapter gives a general definition of "Benchmarking" how it will use it throughout this paper.

@@

While Fixed-Work benchmarks are easy to understand and implement, it introduces an implicit performance bottleneck that is also known as \textit{Amdahl's Law}. It states that the maximum possible performance improvement that can be achieved in Fixed-Work Benchmarks is upper-bounded by the fraction of work that is performed by the component that is being improved (See \cite{Kounev} p. 9f).

@@

	Both of these challenges are heavily influenced by the length of software release cycles and the resulting amount of commits throughout one cycle. As an example one can take a look at a few open-source repositories. The C++ test framework \textit{googletest}\footnote{See \url{https://github.com/google/googletest}} by Google with it's most recent release (\textit{1.11.0}) in June 2021, while the last release before that (\textit{1.10.0}) was released in October 2019. During this period of 28 months roughly 700 commits have been added to the repository. Similarly for the C implementation of the \texttt{lz4} compression algorithm that can be found on GitHub\footnote{\url{https://github.com/lz4/lz4}} in the 15 months between the two most recent releases about 250 commits have been added to the code.

@@

	\subsection{Continuous Benchmarking}
In this section a technique that will further be referred to as \textit{Continuous Benchmarking} will be elaborated further. First the motivation of this approach will be discussed to give a working definition. Secondly a selection of possible continuous benchmarking strategies will be outlined. Lastly the possible benefits and potential challenges will be showcased and weighted to give an assessment of using such a technique.

	\subsubsection{Definition}
	As previously discussed in \autoref{ssec:bench_rel} benchmarking only before releases introduces several challenges. The general motivation of continuous benchmarking is to address these by, simply said, generally adding more data points to evaluate. These added data points can then be used for a more detailed analysis of the systems performance over time. The general idea is similar to those used in \gls{ci} and \gls{cd}. By running tests and releases more frequently, often after every commit, these methods aim to shorten release cycles. \gls{cb} however, is a term and methodology that seems not yet commonly defined in academia.

@@

	\paragraph{}As for any benchmark also \gls{cb} should only be run against a correctly working software. Therefore if integrated into a continuous integration pipeline, \gls{cb} should be among the last steps such that the correct functionality of the software is already ensured by the \gls{cd} tests beforehand (See \cite{grambow2019} p. 2).

	Similar to \gls{cd} approaches in \gls{cb} the question also arises when to run the \gls{cb}. For \gls{cd}, optimally all tests and operations should be performed after every commit. However in practice this might not always be feasible due to the amount of tests and the associated compute time with that which is why complex software systems often already use a multi-tiered approach to \gls{cb}. A similar question arises when deciding on when to run \gls{cb}. Running benchmarks after every commit can pose challenges which will be further evaluated in the following section. However for this paper our main focus will not be on when to run \gls{cb} but rather how the data produced by a \gls{cb} can be used to detect performance changes.

@@

		Using \gls{cb} opens up several opportunities for improvements to the development process of software. First of all it gives a regular overview of a software systems performance. However the usability of these results is also heavily dependent on the quality of the testbed. Especially the use of cloud infrastructure can introduce additional noise to the performance of a system. As has been shown by Daly and Ingo in \cite{daly2019} reducing the noise in such a scenario is a heavily application specific task that can take several weeks of investigating the systems behavior.

	Running performance benchmarks more frequently naturally also enables to detect performance changes more quickly. As a result it also makes it easier to isolate the cause of the change as there may only be fa few commits between the two benchmarks. However again the comparability of these results heavily depends how and if we can take into account the testbed's noise accordingly as to not erroneously flag outliers from noise as actual performance changes. In \cite{daly2021} Daly proposed a technique called \textit{canary testing} to counter these issues. However, in their evaluation canary testing was not viable in their productive system and resulted in more cost than benefit (See \cite{daly2021} p. 6f).

	Another point to consider when using \gls{cb} is the additional data that will inherently get produced. Especially for modern software systems there is not only a single benchmark that is of interest but rather multiple benchmarks with for example different system configurations or workloads. Therefore as Daly et al. and Grambow et al. conclude in their works that manually reviewing the performance data for \gls{cb} is not feasible(See \cite{daly2020} p. 2, \cite{grambow2019} p. 5). As a result, to assist in \gls{cb} and especially in detecting performance changes, automated approaches and toolings are required to pre-process this data and possibly detect performance changes automatically. Approaches for such tools have been described by Daly et al. (\cite{daly2020} p. 2, \cite{daly2021} p. 2ff) and Grambow et al.(\cite{grambow2019}). One of the techniques commonly referred to in this context is change point detection which will be covered in a little more detail in \autoref{sec:cp_detection}.

@@

	\section{Change Point Detection}
	\label{sec:cp_detection}
	This chapter will give a brief introduction into a technique from signal processing known as \textit{Change Point Detection}. It will briefly introduce the term Change point and it's distinction from outliers and introduce one specific change point detection algorithm.

@@

	\begin{figure}[ht!]
		\centering
		\begin{subfigure}[b]{0.45\textwidth}
			\centering
			\includegraphics[width=\textwidth]{outlier}
			\caption{Visualization of an outlier}
		\end{subfigure}
		\begin{subfigure}[b]{0.45\textwidth}
			\centering
			\includegraphics[width=\textwidth]{changepoint}
			\caption{Visualization of a change point}
		\end{subfigure}
		\caption{Visual difference of an outlier and a change point}
		\label{fig:outlier_cp}
	\end{figure}

	In other words, a change point indicates that the overall underlying signal undergoes a noticeable change after that point, while an outlier is only a temporary effect that generally does not persist. \autoref{fig:outlier_cp} shows a visualization of this distinction.

	The challenge of finding such change points is commonly referred to as "Change Point Detection" and is defined by Aminikhanghahi and Cook as "the problem of finding abrupt changes in data when a property of the time series changes" (\cite{Samaneh2016}). In the domain of (continuous) benchmarking, change point detection can be used as a tool to determine points at which a performance regression (or improvement) occurs. With this regard, the performance of a system throughout different commits can be interpreted as a series of time-ordered series, which represents the underlying signal that is analyzed by a change point detection algorithm (See \cite{daly2020} p. 69)

	\subsection{E-Divisive Means}
	\label{ssec:cp_edivisive}
	One algorithm for detecting change points is the E-Divisive Means algorithm. In contrast to other change point detection algorithms it offers a few benefits described by Daly et al. (See \cite{daly2020} p. 69):

	\begin{itemize}
		\item No distributional assumptions required
		\item Can identify multiple change points using recursion
		\item Does not require training data
	\end{itemize}

	The algorithm is based on hierarchically dividing the given time-series into multiple clusters. The recursive approach can be summarized in three steps:
	\begin{enumerate}
		\item For each point in the time series calculate a statistic $\hat{Q}$
		\item Select the largest $\hat{Q}$ as the next change point
		\item This divides the time series into two new clusters
	\end{enumerate}

@@

	To perform the experiment as described in \autoref{ssec:exp_goals} the open-source implementation \texttt{lz4}\footnote{\url{https://github.com/lz4/lz4}} will be used. \texttt{lz4} is a compression library with an active commit history. In total \texttt{lz4} offers 31 releases and more than 2500 commits in total\footnote{As of 16.09.2021}.

@@

	Lzbench offers the option to disable all the other compression algorithms at compile time. The \texttt{lz4} version that is used for the benchmark can simply be changed by overwriting the corresponding lz4 files within the lzbench repository before compilation. In the experiment a slightly adapted version of Lzbench will be used that has the other compression algorithms disabled at compile time and also incorporates some minor changes that were required to ensure compilation on the cluster infrastructure.

@@

	Since the \texttt{lz4} repository consists of over 2500 commits benchmarking each individual commit is infeasible for the scope of this paper as it would both generate too much data to evaluate properly as well as strain the available shared server resources too much. Therefore a two staged approach will be used:
\begin{itemize}
	\item In the first step, a benchmark will be performed on the publicly available releases\footnote{Based on the tags created in the GitHub repository}. The results of these benchmarks will be used to identify a commit range with a significant performance change
	\item All commits in the identified commit range will be benchmarked for a more detailed and thorough analysis.
\end{itemize}

The benchmark itself will be performed on the cluster infrastructure of the Software Engineering Chair of Saarland University using Slurm\footnote{See \url{https://slurm.schedmd.com/documentation.html}}. The machines that are used are part of the \texttt{kine} cluster with the following hardware configuration\footnote{CPU Information has been acquired using the \texttt{lscpu} command}:

@@

	Based on these results the commit range between \textit{v1.8.1} and \textit{v1.8.2} was selected as the commit range for a more thorough and detailed analysis. The results of this analysis are further discussed in \autoref{sec:exp_evaluation}. However as the data for the lz4 benchmark only indicated a performance improvement and the goal of this work is to evaluate the detection both performance improvements and regressions, the release benchmarks were repeated with the \texttt{lz4hc} compression algorithm. The \texttt{lz4hc} implementation is a variant of the \texttt{lz4} implementation that features a higher compression rate at the cost of lower compression speed. As \texttt{lz4hc} is also included in the same source code as \texttt{lz4} the only adaption required to the benchmark was to replace the compression algorithm used in the command line parameters passed.
